Limited physiological plasticity and reduced relative variance in physiological rates of ectotherm populations under climate change

### 2024-07-01 ms.qmd
Fix in fig legend. Adding code chunk arguments and just cleaning some code to make it easier to read.

```{r, fig-concept}
#| label: fig-concept
#| fig-width: 8
#| fig-height: 12
#| fig-cap: "Calculations of acute and acclimation effect sizes. (A) Two idealised thermal performance curves for animals held at 'cold' ('blue') temperatures and warm ('red') temperatures. Shaded blue and red areas are the thermal optima of the performance curves. Physiological rates are measured for a sample of ectotherms at two different temperatures along the thermal performance curves ($T_{1}$ = 20$^\circ$C and $T_{2} = 30$^\circ$C) for both curves. At each temperature a mean physiological rate (R) (points) and its standard deviation (SD) (error bars above and below mean) are estimated. R1.1 and R1.2 are the rates and associated SD (subscripted) for the cold acclimated animals at temperature 1 and 2, respectively. R2.1 and R2.2 are the rates and associated SD (subscripted) for the warm acclimated animals at temperature 1 and 2, respectively. An example of how acute and acclimation $lnCVR_{Q_{10}}$ are calculated from the treatments within the study is provided on the right-hand side of the figure with reference to each of the four possible groups. Note that, two acute effect sizes can be calculated, one for the cold acclimated animals and one for the warm acclimated animals. (B) Species are expected, *a priori*, to vary in their thermal performance curves (thin lines) around an average (thick black line). We restricted our data to areas of each species' performance curve that fell within the natural thermal range of the species (thick lines on each species-level curve). However, given it was not possible to measure the full performance curve for each species some test temperatures within studies may have converged on or moved past the thermal optima. In such cases, we expected our $Q_{10}$ effect sizes to be smaller as indicated by comparing the black dashed lines to grey dashed lines."

@@

Different trait categories showed different acclimation responses across habitat types (see Supplement; @fig-S2). Acclimation capacity also did not vary by life-history stage and there were no differences between $lnRR_{Q_{10_{acclimation}}}$ and $lnRR_{Q_{10_{acute}}}$ between adult and juveniles (@fig-cvrlh A-C) (Adult-Juvenile (Acute): `r mean(acute_all_lh)`, 95% CI: `r quantile(acute_all_lh, 0.025)[1]` to `r quantile(acute_all_lh, 0.975)[1]`, $p_{MCMC}$ = `r pmcmc(acute_all_lh)`; Adult-Juvenile (Acclimation): `r mean(acclim_all_lh)`, 95% CI: `r quantile(acclim_all_lh, 0.025)[1]` to `r quantile(acclim_all_lh, 0.975)[1]`, $p_{MCMC}$ = `r pmcmc(acclim_all_lh)`).

@@

The total proportion of heterogeneity in $lnCVR_{Q_{10}}$ was lower compared to $lnRR_{Q_{10}}$ ($I^2_{sv}$ = `r mean(tot_i2_cvr)`, 95% CI: `r quantile(tot_i2_cvr, 0.025)` to `r quantile(tot_i2_cvr, 0.975)`), with most variation being driven by between-study and trait differences (see Supplement;  @fig-s3). Each life-history stage exhibited the same pattern of variance change in each of the habitats (@fig-cvrlh). Reduced relative variance was particularly prominent for resting metabolic rates and sprint speed although traits differed in whether they exhibited a reduction in variation in physiological rates at high temperatures (@fig-cvrtrait).

@@

```{r, detail_trait}

## Detailed trait category analyses for lnRR
rerun = FALSE
if(rerun){
#################################
  # Detailed Trait Categories
  # For supplement
#################################

@@

  # More realistic model with traits in marine species. Remove antioxidants because there are just too few effects. Note that, for orchardplot we can only fit one random slope, so we'll just use obs as results are unchanged and Bayesian models have it accounted for
   model1.3_m <- metafor::rma.mv(effect_size ~  max_temp_c + c_accl_time + type + trait_category + type:trait_category, V = V, 
                              random = list(~1 | record_num,
                                            ~1 | species_full,
                                            ~1 + max_temp_c | spp,
                                            ~1 | trait,
                                            ~1 | obs),
                              R = list(species_full = m_phylo), dfs = "contain", 
                              test = "t", 
                              data = lnRRQ10_data_m)

@@

  orchard_plot(model1.3_m_het, mod = "trait_category", at = list(type = c("acclim", "acute")), by = "type", xlab = TeX("$lnRR_{Q_{10}}$"), group = "spp", angle = 0)

@@

  model2.3m <- metafor::rma.mv(effect_size ~  max_temp_c + c_accl_time + type + trait_category + type:trait_category, V = V, 
                              random = list(~1 | record_num,
                                            ~1 | species_full,
                                            ~1 + max_temp_c| spp,
                                            ~1 | trait,
                                            ~1 + trait_category| obs),
                              R = list(species_full = m_phylo_cvr), dfs = "contain", 
                              test = "t", struct = c("GEN", "HCS"),  phi = 0, data = lnCVRQ10_data_m))
  saveRDS(model2.3m, file = "./output/models/lnCVRQ10_model2.3m")

  # NOte that we need to drop interaction for terestrial as it's not able to estimate. 
  model2.3t <- metafor::rma.mv(effect_size ~  max_temp_c + c_accl_time + type + trait_category, V = V, 
                              random = list(~1 | record_num,
                                            ~1 | species_full,
                                            ~1 + max_temp_c | spp,
                                            ~1 | trait,
                                            ~1 + trait_category | obs),
                              R = list(species_full = t_phylo_cvr), dfs = "contain", 
                              test = "t", struct = c("GEN", "HCS"),  phi = 0, data = lnCVRQ10_data_t)
  saveRDS(model2.3t, file = "./output/models/lnCVRQ10_model2.3t")

@@

#### *Plots of $I^2$ for multi-level models*
```{r, fig-s3}
#| label: fig-s3
#| fig-cap: "$I^2$ estimates. A) $lnRR_{Q_{10}}$ B) $lnCVR_{Q_{10}}$ and C) $lnVR_{Q_{10}}$." 
#| fig-align: center

>>>>

```{r, fig-concept}
#| label: fig-concept
#| fig-width: 8
#| fig-height: 12
#| fig-cap: "Calculations of acute and acclimation effect sizes. (A) Two idealised thermal performance curves for animals held at 'cold' ('blue') temperatures and warm ('red') temperatures. Shaded blue and red areas are the thermal optima of the performance curves. Physiological rates are measured for a sample of ectotherms at two different temperatures along the thermal performance curves ($T_{1}$ = 20$^\\circ$C and $T_{2} = 30$^\\circ$C) for both curves. At each temperature a mean physiological rate (R) (points) and its standard deviation (SD) (error bars above and below mean) are estimated. R1.1 and R1.2 are the rates and associated SD (subscripted) for the cold acclimated animals at temperature 1 and 2, respectively. R2.1 and R2.2 are the rates and associated SD (subscripted) for the warm acclimated animals at temperature 1 and 2, respectively. An example of how acute and acclimation $lnCVR_{Q_{10}}$ are calculated from the treatments within the study is provided on the right-hand side of the figure with reference to each of the four possible groups. Note that, two acute effect sizes can be calculated, one for the cold acclimated animals and one for the warm acclimated animals. (B) Species are expected, *a priori*, to vary in their thermal performance curves (thin lines) around an average (thick black line). We restricted our data to areas of each species' performance curve that fell within the natural thermal range of the species (thick lines on each species-level curve). However, given it was not possible to measure the full performance curve for each species some test temperatures within studies may have converged on or moved past the thermal optima. In such cases, we expected our $Q_{10}$ effect sizes to be smaller as indicated by comparing the black dashed lines to grey dashed lines."

@@

Different trait categories showed different acclimation responses across habitat types (@fig-btrait and see Supplement; @fig-S2). Acclimation capacity also did not vary by life-history stage and there were no differences between $lnRR_{Q_{10_{acclimation}}}$ and $lnRR_{Q_{10_{acute}}}$ between adult and juveniles (@fig-cvrlh A-C) (Adult-Juvenile (Acute): `r mean(acute_all_lh)`, 95% CI: `r quantile(acute_all_lh, 0.025)[1]` to `r quantile(acute_all_lh, 0.975)[1]`, $p_{MCMC}$ = `r pmcmc(acute_all_lh)`; Adult-Juvenile (Acclimation): `r mean(acclim_all_lh)`, 95% CI: `r quantile(acclim_all_lh, 0.025)[1]` to `r quantile(acclim_all_lh, 0.975)[1]`, $p_{MCMC}$ = `r pmcmc(acclim_all_lh)`).

@@

The total proportion of heterogeneity in $lnCVR_{Q_{10}}$ was lower compared to $lnRR_{Q_{10}}$ ($I^2_{sv}$ = `r mean(tot_i2_cvr)`, 95% CI: `r quantile(tot_i2_cvr, 0.025)` to `r quantile(tot_i2_cvr, 0.975)`), with most variation being driven by between-study and trait differences (see Supplement;  @fig-s3). Each life-history stage exhibited the same pattern of variance change in each of the habitats (@fig-cvrlh). Reduced relative variance was particularly prominent for resting metabolic rates and sprint speed although traits differed in whether they exhibited a reduction in variation in physiological rates at high temperatures (@fig-btrait).

@@

```{r, detail_trait}
#| label: detail_trait
#| echo: false
#| include: false
#| warning: false

## Detailed trait category analyses for lnRR

@@

rerun = FALSE
if(rerun){
#################################
  # Detailed Trait Categories
  # For supplement
#################################
  # More realistic model with traits in marine species. Remove antioxidants because there are just too few effects. Note that, for orchardplot we can only fit one random slope, so we'll just use obs as results are unchanged and Bayesian models have it accounted for
   model1.3_m <- metafor::rma.mv(effect_size ~  max_temp_c + c_accl_time + type + trait_category + type:trait_category, V = V, 
                              random = list(~1 | record_num,
                                            ~1 | species_full,
                                            ~1 + max_temp_c | spp,
                                            ~1 | trait,
                                            ~1 | obs),
                              R = list(species_full = m_phylo), dfs = "contain", 
                              test = "t", 
                              data = lnRRQ10_data_m)

@@

@@

  model2.3m <- metafor::rma.mv(effect_size ~  max_temp_c + c_accl_time + type + trait_category + type:trait_category, V = V, 
                              random = list(~1 | record_num,
                                            ~1 | species_full,
                                            ~1 + max_temp_c| spp,
                                            ~1 | trait,
                                            ~1 + trait_category| obs),
                              R = list(species_full = m_phylo_cvr), dfs = "contain", 
                              test = "t", struct = c("GEN", "HCS"),  phi = 0, data = lnCVRQ10_data_m)
  saveRDS(model2.3m, file = "./output/models/lnCVRQ10_model2.3m")

  # Note that we need to drop interaction for terrestrial as it's not able to estimate. 
  model2.3t <- metafor::rma.mv(effect_size ~  max_temp_c + c_accl_time + type + trait_category, V = V, 
                              random = list(~1 | record_num,
                                            ~1 | species_full,
                                            ~1 + max_temp_c | spp,
                                            ~1 | trait,
                                            ~1 + trait_category | obs),
                              R = list(species_full = t_phylo_cvr), dfs = "contain", 
                              test = "t", struct = c("GEN", "HCS"),  phi = 0, data = lnCVRQ10_data_t)
  saveRDS(model2.3t, file = "./output/models/lnCVRQ10_model2.3t")

@@

#### *Plots of $I^2$ for multilevel models*
```{r, fig-s3}
#| label: fig-s3
#| fig-cap: "$I^2$ estimates. A) $lnRR_{Q_{10}}$ B) $lnCVR_{Q_{10}}$ and C) $lnVR_{Q_{10}}$." 
#| fig-align: center
